---
title: "Workpackage 1-Nonfermentors"
date: "12/14/2022"
format: 
  html:
    code-fold: true
    number-sections: true
execute:
  echo: true
  warning: false
  freeze: false
bibliography: references.bib
csl: nature.csl
---

## Workpackage 1A: <br>Quantitative Relationship of Tpos with Numbers of Viable Gram-Negative Non-fermentor bacteria

### Acinetbacter baumannii
#### Methods

In this package we will establish the quantitative relationship of Tpos and *Acinetobacter baumanii* and *Pseudomonas seruginosa* inoculum in blood culture bottles. <br> <br>

The methods for the assay are a modification of those originally proposed by Kaltsas et al.[@kaltsas2005] Detailed procedures can be found at this [link](Protocols.html). Methodology for preparing the test inoculum was adapted from [CLSI M21A](CLSI_M21A.pdf) and [M26A](CLSI_M26A.pdf) guidelines.[@Clinical_Laboratory_Standards_Institute1999-iy; @Clinical_Laboratory_Standards_Institute1999-iya]

-   Briefly, tubes containing 1.8 mL of pooled healthy human serum are inoculated with 0.2 mL of a series of ten-fold dilutions (5x10^1^ to 5x10^7^ CFU/mL) of the standardised inoculum of each test indicator strains.
-   The sera were are then transferred into BacT⁄ALERT bottles ***without antibiotic inactivating matrix*** (Biomérieux Inc) for aerobic incubation for 24 hours and monitored for time to positivity.

#### Results

The relationship between Tpos and the *A. baumanii* inoculum is shown in [@fig-1f] a linear relationship was observed from approximately 10^1^-10^8^ *K. pneumoniae* CFU/mL and a Tpos measured from 10.5 hours-4.5 hours over the tested inoculum range, with R^2^ of 0.92-0.94. The linear relationship of Tpos versus inoculum was consistent if the inoculum was prepared in PBS or pooled human serum as shown in [@tbl-1f].

::: callout-note
In the paper by Kaltsas et al. [@kaltsas2005] the reported inoculum was the inoculum introduced in to the bloodculture bottle, 30-40 mL of growth media (depending on the manufacturer).

Therefore we have followed this precedent of reporting the inoculum introduced into each bottle. Therefore, this does not represent the final test inoculum that need to account for the total 42 mL volume in each bloodculture bottle.

All experiments were performed using bloodculture bottles ***without antibiotic inactivating matrix***
:::

```{r}
#| label: fig-1f
#| fig-cap: "Relationship of time-to-positivity (Tpos) versus test inoculum for Acinetobacter baumanii"


library (ggplot2)
library(scales)
theme_set(theme_bw())
## import raw data from .csv file
library (readxl)
abaumani <- read_excel("abaum_pbs.xlsx")
## plot raw data as x-y graph Tpos graph vs. drug concentrations stratified by dilution matrix, method="lm" is the method for linear regression
fig1 <-ggplot(abaumani, aes(x=inoculum, y=tpos)) + geom_point(size=4, alpha = 0.5) + 
scale_y_continuous(name="Tpos (hr)", limits=c(4,12)) +
theme(legend.text=element_text(size=12)) +
geom_smooth(aes(linetype=diluent), method=lm , color="black", fill="#69b3a2", se=TRUE, inherit.aes = TRUE )
fig1 + theme_bw(base_size = 14)+ scale_x_log10(name="Inoculum CFU/mL", breaks = trans_breaks("log10",n=7, function(x) 10^x),labels = trans_format("log10", math_format(10^.x)))

```

```{r}
#| label: tbl-1f
#| tbl-cap: Effect of inoculum diluent on Tpos for Acinetobacter baumanii
library(forestmangr)
library(kableExtra)
library (readxl)
abaumani <- read_excel("abaum_pbs.xlsx")
df <- data.frame (abaumani)
table1<-lm_table(df, log(inoculum) ~ tpos, "diluent")
kbl(table1)%>%
kable_paper("hover", full_width = F, position="left")
```

::: callout-note
These data confirm that Tpos exhibits a wide dynamic range as a surrogate indicator for viable CFU/mL.
:::

------------------------------------------------------------------------
### Pseudomonas aeruginosa

```{r}
#| label: fig-2f
#| fig-cap: "Relationship of time-to-positivity (Tpos) versus test inoculum for Pseudomonas aeruginosa"

library (ggplot2)
library(scales)
theme_set(theme_bw())
## import raw data from .csv file
library (readxl)
pseudo <- read_excel("pseudo_pbs.xlsx")
## plot raw data as x-y graph Tpos graph vs. drug concentrations stratified by dilution matrix, method="lm" is the method for linear regression
fig2f <-ggplot(pseudo, aes(x=inoculum, y=tpos, color=isolates, fill=isolates)) + geom_point(size=4, alpha = 0.5) + 
scale_y_continuous(name="Tpos (hr)", limits=c(4,12)) +
theme(legend.text=element_text(size=12)) +
geom_smooth(aes(linetype=diluent), method=lm , color="black", fill="#69b3a2", se=TRUE, inherit.aes = TRUE )
fig2f + theme_bw(base_size = 14)+ scale_x_log10(name="Inoculum CFU/mL", breaks = trans_breaks("log10",n=7, function(x) 10^x),labels = trans_format("log10", math_format(10^.x)))


```

```{r}
#| label: tbl-2f
#| tbl-cap: Effect of inoculum diluent on Tpos for Acinetobacter baumanii
library(forestmangr)
library(kableExtra)
library (readxl)
pseudo <- read_excel("pseudo_pbs.xlsx")
df2 <- data.frame (pseudo)
table2<-lm_table(df, log(inoculum) ~ tpos, "diluent")
kbl(table1)%>%
kable_paper("hover", full_width = F, position="left")
```

::: callout-note
These data confirm that Tpos exhibits a wide dynamic range as a surrogate indicator for viable CFU/mL.
:::
