---
title: "Workpackage 1"
execute:
  echo: false
  warning: false
bibliography: references.bib
csl: nature.csl
---

## Workpackage 1A

Establishing the quantitative relationship of TPOS and KPC-*Klebsiella pneumoniae* inoculum in blood culture bottles. A modification of the methods proposed by Kaltsas et al.[@kaltsas2005]. Detailed methods can be found following this [link](Protocols.html)

Briefly, tubes containing 1.8 mL of pooled healthy human serum were inoculated with 0.2 mL of a series of seven ten-fold dilutions (5x10^1^ to 5x10^7^ CFU/mL) of the standardised inoculum of each test indicator strains listed in Table 1. The inoculated sera are then transferred into BacT⁄ALERT bottles without inactivating matrix (Biomérieux Inc) for aerobic incubation and monitored for time to positivity. Tpos results will be used to establish preliminary assay quality control ranges and confirm the utility of isolates as selected "indicator" organisms and quality control ranges at specific inoculum thresholds. Experiments will be performed in at least 10 replicates. The main goal of this workpackage is to demonstrate that Tpos results under the assay conditions exhibit a reproducible and reliable relationship with bacterial inoculum for representative KPC-carbapenemase producing strains of *K. pneumoniae.*

Experiments were repeated in triplicate for five KPC-carbapenemase producing *K. pneumoniae* strains (KPC_KPCA,B,C and NDM, and VIM producing strains) and a reference *K. pneumoniae* ATCC strain producing ESBL only. Detailed information on the isolates can be found [here](Protocols.html). We also compared how inoculation of the organism suspension prepared in phosphate buffered saline (pbs-0.9%) versus pooled human serum (serum) affected Tpos.

[Figure 1]. shows the relationship between Tpos and the *K. pneumoniae* inoculum fitted by Loess. A linear relationship was evident from approximately 10^1^-10^8^ *K. pneumoniae* CFU/mL and a Tpos measured in the BACTEC-Alert culture system from 10.5 hours-4.5 hours over the tested inoculum range. *These data suggest Tpos has a high dynamic range as a possible biomarker*.

::: {style="red"}
Note*: we need to add a special note about the actual inoculum in the bottle*
:::

#### Figure 1

```{r}
library (ggplot2)
theme_set(theme_bw())
wp1 <- read.csv("~/Desktop/ACUTE/wp1a.csv")
fig1 <-ggplot(wp1, aes(x=inoculum, y=tpos, color=isolates, shape=diluent, fill=isolates)) + geom_point(size=4, alpha = 0.5) + 
scale_x_log10 () +
scale_y_continuous(name="Time to positivity (Tpos)", limits=c(4,12)) +
ggtitle("Relationship between Tpos and inoculum")+
theme(legend.text=element_text(size=12)) +
geom_smooth(aes(linetype=diluent), method=lm , color="black", fill="#69b3a2", se=TRUE, inherit.aes = TRUE )
fig1 + theme_bw(base_size = 16)

```

<br> 



#### Figure 2. 

```{r}
library (ggplot2)
theme_set(theme_bw())
ceftaz <- read.csv("~/Desktop/ACUTE/ceftazidime.csv")
ggplot(ceftaz, aes(x=conc, y=tpos, color=isolate, shape=isolate)) + geom_point(size=4, alpha = 0.7)+ 
scale_x_log10(name="Ceftazidime/Avibactam Concentration, mg/L") +
scale_y_continuous(name="Time to positivity (Tpos)", limits=c(4,25)) 

```

#### Figure 3. Relationship of ceftazidime-avibactam concentrations vs. time to positivity for KP_KPC_B

```{r}
# | code-fold: true
library(drda)
wp2 <- read.csv("~/Desktop/ACUTE/ec50_KPCB.csv")
fit <- drda(tpos ~ conc, data=wp2, mean_function = "logistic4", max_iter = 1000)
plot(fit, xlab = "log(conc)", ylab = "Tpos hrs")

```

```{r}
summary (fit)
```
