---
title: "Workpackage 1"
date: "9/23/2022"
format: 
  html:
    code-fold: true
execute:
  echo: true
  warning: false
bibliography: references.bib
csl: nature.csl
---

## Package 1A

**Establishing the quantitative relationship of TPOS and KPC-*Klebsiella pneumoniae* inoculum in blood culture bottles.** <br> A modification of the methods proposed by Kaltsas et al.[@kaltsas2005] Detailed methods can be found following this [link](Protocols.html). Methodology for preparing the test inoculum was adapted from [CLSI M21A](CLSI_M21A.pdf) and [M26A](CLSI_M26A.pdf) guidelines.[@Clinical_Laboratory_Standards_Institute1999-iy;@Clinical_Laboratory_Standards_Institute1999-iya]

Briefly, tubes containing 1.8 mL of pooled healthy human serum were inoculated with 0.2 mL of a series of seven ten-fold dilutions (5x10^1^ to 5x10^7^ CFU/mL) of the standardised inoculum of each test indicator strains. The inoculated sera were then transferred into BacT⁄ALERT bottles without inactivating matrix (Biomérieux Inc) for aerobic incubation and monitored for time to positivity. Tpos results were used to establish preliminary assay quality control ranges by testing in triplicate for five KPC-carbapenemase producing *K. pneumoniae* strains (KPC_KPCA,B,C and NDM, and VIM producing strains) and a reference *K. pneumoniae* ATCC strain producing ESBL only. Detailed information on the isolates can be found [here](Protocols.html). We also compared how inoculation of the organism suspension prepared in phosphate buffered saline (pbs-0.9%) versus pooled human serum (serum) affected Tpos.

Figure 1 shows the relationship between Tpos and the *K. pneumoniae* inoculum. A linear relationship was observed from approximately 10^1^-10^8^ *K. pneumoniae* CFU/mL and a Tpos measured from 10.5 hours-4.5 hours over the tested inoculum range, with R^2^ of 0.92-0.94. *These data confirm thatTpos has a wide dynamic range as a surrogate indicator for viable CFU/mL.*

::: callout-note
We need to add an explanation about the *actual* inoculum in the bottles
:::

### Figure 1. Tpos vs. test inoculum

```{r}
library (ggplot2)
library(scales)
theme_set(theme_bw())
wp1 <- read.csv("~/Desktop/ACUTEWEBSITE/wp1a.csv")
fig1 <-ggplot(wp1, aes(x=inoculum, y=tpos, color=isolates, shape=diluent, fill=isolates)) + geom_point(size=4, alpha = 0.5) + 
scale_y_continuous(name="Tpos (hr)", limits=c(4,12)) +
theme(legend.text=element_text(size=12)) +
geom_smooth(aes(linetype=diluent), method=lm , color="black", fill="#69b3a2", se=TRUE, inherit.aes = TRUE )
fig1 + theme_bw(base_size = 14)+ scale_x_log10(name="Inoculum CFU/mL", breaks = trans_breaks("log10",n=7, function(x) 10^x),labels = trans_format("log10", math_format(10^.x)))

```

```{r}
library(forestmangr)
library(kableExtra)
wp1 <- read.csv("~/Desktop/ACUTEWEBSITE/wp1a.csv")
df <- data.frame (wp1)
table1<-lm_table(df, log(inoculum) ~ tpos, "diluent")
kbl(table1)%>%
kable_paper("hover", full_width = F, position="left")
```

## Package 1B

This workpackage explores how antibiotic concentrations impact Tpos. <br>

The effect of increasing fixed ceftazidime/avibactam (CAZ/AVI) concentrations (x:x) on Tpos was investigated at an inoculum of 1x10^4^ CFU/mL of *K. pneumoniae* KPC_B (MIC X mg/L) and *K. pneumonia* NDM producing isolate (negative control, MIC X mg/L ). All experiments were performed in triplicate with an incubation period of 24 hours. Figure 2 shows a marked increased in the Tpos for KPC_B from \< 10 hrs to [\>]{.underline} 24 hrs when CAZ/AVI concentrations surpassed 1 mg/L. In contrast, the negative control KPC_NDM strain exhibited consistent Tpos \< 10hr at all test concentrations consistent with lack of antimicrobial activity.

### Figure 2. Impact of CAZ/AVI concentrations on Tpos

```{r}
library (ggplot2)
theme_set(theme_bw())
ceftaz <- read.csv("~/Desktop/ACUTE/ceftazidime.csv")
ggplot(ceftaz, aes(x=conc, y=tpos, color=isolate, shape=isolate)) + geom_point(size=4, alpha = 0.7)+ 
scale_x_log10(name="CAZ/AVI, mg/L") +
scale_y_continuous(name="Tpos (hr)", limits=c(4,25)) 

```

<br>

A four-parameter logistic regression model was fit to the experimental data (Figure 3) for the KPC_B isolate (Figure 3). The model estimated an EC~50~ for the total Tpos effect at CAZ/AVI 2.15 mg/L (95% CI 2.04-2.27). The EC~90~ was estimated at 2.75 mg/L (95% CI 2.21-3.30).

<br>

### Figure 3. In vitro pharmacodynamics of CAZ/AVI concentrations vs. Tpos

```{r}
# | code-fold: true
library(drda)
wp2 <- read.csv("~/Desktop/ACUTE/ec50_KPCB.csv")
fit <- drda(tpos ~ conc, data=wp2, mean_function = "loglogistic4", max_iter = 1000)
plot(fit, xlab = "CAZ/AVI (conc)", ylab = "Tpos (hr)")

```

```{r}
library(drda)
library(broom)
library(kableExtra)
wp2 <- read.csv("~/Desktop/ACUTEWEBSITE/ec50_KPCB.csv")
fit <- drda(tpos ~ conc, data=wp2, mean_function = "logistic4", max_iter = 1000)
ed<-effective_dose(fit, y = c(0.10,0.25,0.50,0.75,0.90,0.95))
kbl(ed)%>%
kable_paper("hover", full_width = F, position = "left")
```

<br>

Ceftazidime/avibactam was tested as a combination prepared in pure powders against the KPC_B and ATCC strains. The results of data fitted with a 4-parameter logistic model are shown in Figure 4.

## Figure 4. In vitro pharmacodynamics of ceftazidime/avibactam, 19-07-2022 

::: callout-important
Need to explain why this experiment failed
:::

```{r}
# | code-fold: true
library(drda)
cazavi1 <- read.csv("~/Desktop/ACUTEWEBSITE/19_07_cazavi.csv")
fit <- drda(tpos ~ caz, data=cazavi1, mean_function = "loglogistic4", max_iter = 1000)
plot(fit, xlab = "CAZ/AVI (conc)", ylab = "Tpos (hr)")

```
